---
title: "Assignment 1"
output: html_document
---
  
A. Loading and pre-processing the data in R  
  
```{r, echo=TRUE}
  
# R script Assignment Activity monitoring data #  

# loading and preprocessing the data # 
activity <- read.csv("C:\\D\\Weiterbildung\\ReproducibleResearch\\repdata_data_activity\\activity.csv")

is.data.frame(activity)

head(activity)

```

B. What is mean total number of steps taken per day?

In this part, we ignore the missing values in the dataset. 

1. The total number of steps taken per day:

```{r, echo=TRUE}
# What is mean total number of steps taken per day # 
dim(activity)
activity.comp <- activity[!is.na(activity$steps),]
dim(activity.comp)
head(activity.comp)
length(unique(activity.comp$date))

dates <- unique(activity.comp$date)

# create an empty vector steps.day # 
steps.day <- rep(x=NA, times=length(dates))

# calculate the total number of steps taken per day # 
for(i in 1:length(dates)){
  steps.day[i] <- sum(activity.comp$steps[activity.comp$date == dates[i]])
  
}
steps.day

```

2. Histogram of the total number of steps taken each day:


```{r, echo=TRUE}

# plot a histogram of the total number of steps per day #
hist(steps.day)
#     , breaks=length(dates))
# axis(1, labels = substr(dates, start=6, stop=10)[c(1,11,21,31,41)]  )
box()

substr(dates, start=6, stop=10)

```
3. Mean and Median of the total number of steps taken per day.

```{r, echo=TRUE}

# mean of the total number of steps taken per day

# create an empty vector steps.day # 
mean.steps.day <- rep(x=NA, times=length(dates))

# calculate the total number of steps taken per day # 
for(i in 1:length(dates)){
  mean.steps.day[i] <- mean(activity.comp$steps[activity.comp$date == dates[i]])
  
}
mean.steps.day

# median of the total number of steps taken per day

# create an empty vector steps.day # 
median.steps.day <- rep(x=NA, times=length(dates))

# calculate the total number of steps taken per day # 
for(i in 1:length(dates)) {
  median.steps.day[i] <- median(activity.comp$steps[activity.comp$date == dates[i]])
  
}
median.steps.day

```

C. What is the average daily activity pattern?

1. Time series plot of the 5 minute interval and the average number of steps taken, averaged across all days.

```{r, echo=TRUE}
# average daily activity pattern

intervals <- unique(activity.comp$interval)


# create an empty vector steps.day # 
mean.steps.intervals <- rep(x=NA, times=length(intervals))

# calculate the mean number of steps taken per day # 
for(j in 1:length(intervals)){
  mean.steps.intervals[j] <- mean(activity.comp$steps[activity.comp$interval == intervals[j]])
  
}
mean.steps.intervals

plot(x=intervals, y = mean.steps.intervals, type="l")

```

2. Which 5 minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r, echo=TRUE}

# Which 5 minute interval is the largest

max(mean.steps.intervals)

intervals[which(mean.steps.intervals == max(mean.steps.intervals))]

```

D. Imputing missing values

1. The number of missing values in the dataset:

```{r, echo=TRUE}
# Impute missing values

# number of missing values 

length(activity$steps)
sum(is.na(activity$steps))
sum(is.na(activity$date))
sum(is.na(activity$interval))

```

2. Strategy for filling in the data. 

As an attempt to impute the missing data, I will use the mean number of steps for 5 minutes interval.

```{r, echo=TRUE}

plot(x=intervals, y = mean.steps.intervals, type="l")


```

3. Create new dataset with the missing data imputed.


```{r, echo=TRUE}

activity.imputed <- activity

# activity.imputed$steps

```
4. Make a histogram of total number of steps taken each day, and calculate and report mean and median total number of steps per day. 

```{r, echo=TRUE}


```

Do these values differ from the estimates from the first part of the assignment?

```{r, echo=TRUE}


```

What is the impact of imputing the missing data on the estimates?


```{r, echo=TRUE}


```

E. Are there differences in the activity patterns between weekdays and weekend?

1. Create a new factor variables in the dataset with two levels, "Weekday" and "weekend".


```{r, echo=TRUE}
# use the dataset with the filled in missing values

weekday <- weekdays(as.Date(activity.imputed$date))

weekend <- rep("weekday", times = 17568 )
weekend[weekdays = "Saturday"] <- "weekend"
weekend[weekdays = "Sunday"]   <- "weekend"
weekend <- as.factor(weekend)

```

2. Make a panel plot containing a time series plot of the 5 minute interval and average number of steps, averaged across all weekdays or weekends.

```{r, echo=TRUE}

library(lattice)

intervals <- unique(activity.imputed$interval)

# create an empty vector steps.day # 
mean.steps.intervals.weekday <- rep(x=NA, times=length(intervals))
mean.steps.intervals.weekend <- rep(x=NA, times=length(intervals))

# calculate the mean number of steps taken per day # 
for(j in 1:length(intervals)){
  mean.steps.intervals.weekday[j] <- mean(activity$steps[activity.imputed$interval == intervals[j] & weekend == "weekday"])
  
}

for(j in 1:length(intervals)){
  mean.steps.intervals.weekend[j] <- mean(activity$steps[activity.imputed$interval == intervals[j] & weekend == "weekend"])
  
}

xyplot(intervals ~ mean.steps.intervals | weekend, data = activity.imputed, type = "l")




```
